---
title: "The GRASS Reporting Card: two-column tier labels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The GRASS Reporting Card: two-column tier labels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 110)
library(grass)
```

*Preview vignette for the §8 Reporting Card of the merged GRASS binary-
rater-reliability paper. The two functions exercised below — `check_asymmetry()`
and `classify()` — ship with `grass` 0.2.x. Tier definitions follow
`paper2/review/framework_notes.md` §0.6 (Column A, three-tier model safety)
and §0.6.5 ("Labels without re-creating Landis-Koch"; Column B, five-tier
use-case fit).*

## What a Reporting Card row looks like

Every panel reported under the GRASS framework carries **two orthogonal
tier columns** alongside the arithmetic:

- **Column A (model safety)** — is `q̂` trustworthy as the primary summary?
  Assigned by `check_asymmetry()` from per-rater `(Ŝe_j, Ŝp_j)` estimates:
  `ok` (δ̂ < 0.05), `caution` (0.05–0.10), or `unsafe` (≥ 0.10). At Tier 3
  the Se = Sp diagonal is visibly broken and the panel needs a latent-class
  fit rather than a single `q̂`.
- **Column B (use-case fit)** — does the panel meet its *declared* use
  case? Assigned by `classify()` from the 95% CI on **EMR_panel** (the
  expected rate at which the panel's majority-vote decision disagrees with
  the latent class) against a tolerance `T` drawn from the use-case ladder.
  Five tiers: `Fails`, `Indeterminate`, `Meets`, `Met with distinction`
  (upper CI below `T/2`), `Exceeds` (upper CI below the next harder use
  case's tolerance).

The columns are independent: a panel can be `ok + Met with distinction`
(the common case), `caution + Meets` (asymmetry-flagged but still passes),
or `unsafe + Exceeds` (latent-class fit needed, but the panel already
clears a harder use case once the fit is trusted).

## Worked example 1: a screening panel, symmetric raters

Five radiologists independently score the same 50 chest films for a
binary screening-eligibility endpoint. A simulation-calibrated latent-
class fit gives per-rater sensitivity and specificity:

```{r}
se <- c(R1 = 0.88, R2 = 0.86, R3 = 0.89, R4 = 0.87, R5 = 0.85)
sp <- c(R1 = 0.86, R2 = 0.87, R3 = 0.88, R4 = 0.85, R5 = 0.86)
```

### Column A: is the Se = Sp diagonal defensible?

```{r}
safety <- check_asymmetry(se = se, sp = sp, rater = names(se))
safety
```

The largest per-rater gap is 0.03, below the `caution` threshold of 0.05,
so the panel sits at **Tier 1 (`ok`)**. We may report `q̂ ± SE` plus EDR and
EMR_panel without per-rater disclosure; PABAK's prevalence-flatness
assumption holds.

### Column B: does the panel meet the screening use case?

Under Tier 1 the panel's operating quality is `q̂ ≈ 0.87` with SE ≈ 0.02
(Wilson-logit); the panel-level misclassification rate under 5-rater
majority vote and its 95% CI are derived from `q̂` and `k`:

```{r}
q_hat <- 0.87
emr_point <- emr_panel(q = q_hat, k = 5)
# Rough 95% CI on EMR via propagating q_hat ± 1.96 * 0.02 through emr_panel()
emr_lo <- emr_panel(q = q_hat + 1.96 * 0.02, k = 5)  # lower EMR = upper q
emr_hi <- emr_panel(q = q_hat - 1.96 * 0.02, k = 5)  # upper EMR = lower q
c(point = emr_point, lower = emr_lo, upper = emr_hi)
```

Hold the panel to the declared screening tolerance `T = 0.10`:

```{r}
verdict <- classify(emr_upper = emr_hi, emr_lower = emr_lo,
                    emr_point = emr_point,
                    use_case = "screening",
                    regime = safety)
verdict
```

`Exceeds`: the upper CI on EMR_panel (`r sprintf("%.3f", emr_hi)`) lies
below the diagnostic tolerance `T_next = 0.05`, so the panel would also
clear the next step up the ladder. `Exceeds` is the only tier that
licenses *promoting* the panel's declared use case — the label points to
a specific operational consequence, not an adjective.

### ### The citation sentence

Every label must appear with its arithmetic — that is what keeps the
label system citable without repeating the Landis-Koch error of
stipulative, regime-free cutoffs:

```{r citation, results = "asis", echo = FALSE}
cat(sprintf(paste0(
  "> *\"The rater panel exceeded the GRASS screening tier ",
  "(q_hat = 0.87 +/- 0.02; EMR_panel = %.3f at k = 5 majority vote, ",
  "95%% CI [%.3f, %.3f]; declared tolerance T = 0.10, ",
  "margin %.3f; panel would also clear the diagnostic tolerance ",
  "T_next = 0.05; inter-rater axis; N = 50; ",
  "Tier 1 model safety per check_asymmetry() delta_hat = %.3f).\"*"),
  emr_point, emr_lo, emr_hi, verdict$margin, safety$delta_hat))
```

Reads like a qualitative verdict; every adjective is reconstructible.

## Worked example 2: smaller pilot, flagged asymmetry, harder use case

Now suppose the same five readers have been re-fielded for a *diagnostic*
use case (tolerance `T = 0.05`). Two changes: the calibration pilot is
only `N = 20` films (so SE on `q̂` is roughly doubled to `0.04`), and
one rater has drifted into a visible Se-favoring asymmetry (a known
quirk of the reading protocol when on shift > 10 hours).

```{r}
se2 <- c(R1 = 0.88, R2 = 0.86, R3 = 0.93, R4 = 0.87, R5 = 0.85)
sp2 <- c(R1 = 0.86, R2 = 0.87, R3 = 0.84, R4 = 0.85, R5 = 0.86)

safety2 <- check_asymmetry(se = se2, sp = sp2, rater = names(se2))
safety2
```

R3's gap of 0.09 is below the `unsafe` threshold but above `caution`, so
the panel sits at **Tier 2**. `q̂` remains reportable as the primary
summary but must carry a caution flag and the `δ̂` value; R3's asymmetry
is disclosed so readers can weigh it.

```{r}
# Wider CI from the smaller pilot
emr_lo2 <- emr_panel(q = q_hat + 1.96 * 0.04, k = 5)
emr_hi2 <- emr_panel(q = q_hat - 1.96 * 0.04, k = 5)
c(lower = emr_lo2, upper = emr_hi2)

verdict2 <- classify(emr_upper = emr_hi2, emr_lower = emr_lo2,
                     emr_point = emr_point,
                     use_case = "diagnostic",
                     regime = safety2)
verdict2
```

`Indeterminate`: the 95% CI on EMR_panel straddles `T = 0.05` — the
point estimate is good, but the pilot is too thin to resolve whether the
panel is below or above the diagnostic tolerance. The operational action
is **not** "fail" — it is "collect more data". Either more subjects
(tighten the CI) or more raters (raise k, drop EMR_panel). The same
panel that `Exceeds` the screening tolerance can be honestly called
`Indeterminate` for diagnostic use; that is exactly the regime-conditioned
behaviour the Landis-Koch scheme cannot express.

## Assembling a two-row Reporting Card

Both verdicts bind cleanly into a single results table:

```{r}
card <- rbind(
  cbind(row_label = "Screening (N = 50)", as.data.frame(verdict)),
  cbind(row_label = "Diagnostic pilot (N = 20)", as.data.frame(verdict2))
)
card[, c("row_label", "use_case", "tolerance", "emr_upper",
         "margin", "tier", "column_a_tier", "column_a_regime")]
```

That data.frame is the literal object the §8 Reporting Card renders in
the paper.

## What the framework commits to

1. **Labels live on EMR_panel**, not on any individual agreement
   coefficient. κ / PABAK / AC1 / α / ICC all feed `q̂` upstream;
   `q̂` and `k` feed EMR_panel; only EMR_panel's CI against `T` assigns
   the use-case tier.
2. **T is declared, not stipulated**, and comes from use-case precedent
   (see [`grass_use_case_ladder()`](../reference/grass_use_case_ladder.html))
   or from the practitioner with explicit disclosure.
3. **"With distinction" means margin within the declared use case**;
   **"Exceeds" means the panel would pass a harder use case**. The
   distinction is load-bearing, not cosmetic — only `Exceeds` licenses
   promoting the panel one step up the ladder.
4. **Every label appears with its arithmetic.** A reviewer who asks
   *"with distinction under what conditions?"* finds the answer in the
   parenthetical.

## Further reading

- `paper2/review/framework_notes.md` §0.6, §0.6.5 — the argument for why
  a label system that sits on EMR_panel with declared tolerances does
  not re-create Landis-Koch (1977) or Koo-Li (2016).
- `paper2/review/merged_paper_scope.md` — §8 Reporting Card in the full
  paper scope.
- `vignette("grass-intro")` — the Column-A-free inter-rater workflow
  that the Reporting Card slots on top of.
